
---
title: "Final Project"
author: "Jaji Pamarthi"
date: "`r Sys.Date()`"
description: "Final Project"
output: distill::distill_article
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE, cache=TRUE)
```

## Bicyclist Accidents Data

### Introduction

Massachusetts is one of the most bicycle friendly city in the USA. There are several laws for the bicyclists and drivers of other means of transport to ensure the safety of the bicyclists. Even after these laws, there are nearly 1,000 bicyclists die and over 130,000 are injured in crashes that occur on roads in the United States every year according to [CDC](cdc.gov/injury/wisqar.) Understanding the data from past accident is important to improve bicyclist safety. This is an interesting dataset for me as I'm currently working on Bicycle Simulator project. Understanding of real-time data like this dataset would help in better design and developing experiments to understand bicyclist behavior as well.



### Dataset Information

This data set is about the bicycle accidents in Britain from year 1979 to 2018.

This Dataset contains data such as the accident index, number of vehicles involved, number of casualties, date and time of accident, speed limit, road and weather conditions, day of accident and finally the road type in which the accident took place in Accidents.csv file. It also includes the gender of person driving the bicycle, severity of the accident and the age group range of the victims in bikers.csv file.

The dataset used is from kaggle and the link to data is [Bike Accidents in Great Britain](https://www.kaggle.com/datasets/johnharshith/bicycle-accidents-in-great-britain-1979-to-2018?select=Accidents.csv)

## Required Libraries

```{r}
library(readr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(wesanderson)
library(corrplot)
library(rmarkdown)
library(stringr)
```

## Read Bikers information and Accident Data  

### Reading csv file and combining the two files

Ignoring that data where is no match found for Accident_Index in Bikers_data to Accident_Index in Accident_data.

```{r}
Accident_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/Data/Accidents.csv")
Bikers_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/Data/Bikers.csv")


joined_data <- merge(Accident_data, Bikers_data, by.x = "Accident_Index", 
             by.y = "Accident_Index", all.x = TRUE, all.y = FALSE) # merging two csv files

paged_table(head(joined_data,20))



```

### Understanding Data 


```{r}

dim(joined_data)

joined_data %>%
  #Unselecting columns with known data or large unique values.
  select(-Accident_Index,-Time,-Date,-Day) %>% 
  lapply(unique) #Finding unique values in each column
```

### Tidy Data Generation


```{r}

# Generating Month name and Year from Data column
joined_data<-joined_data%>%
 mutate(Year_data= format(Date, format="%Y"),Month_number= format(Date, format="%m"))%>%
 mutate(Month_name = month.abb[as.numeric(Month_number)])


#Removing the columns to reduce the data for analysis
tidy_data<-select(joined_data, -Accident_Index, -Time, -Date, -Day,-Month_number)
tidy_data<-select(tidy_data,11,12,10,1:9)

# Arranging the data based on Year followed by name of the month
tidy_data<-tidy_data %>%
  arrange(Year_data,Month_name)



```


## Visualizations


### Research Question 1: Does younger age group have higher accident counts?

```{r}

ggplot(tidy_data, aes(x=Age_Grp)) + 
  geom_bar(fill="bisque4") + xlab("Age range in years")+ylab("Number of Accidents")+
  labs(title = "Fig 1: Count of Accidents based on Age Group")



```

Figure 1. Age group 11-15 years had higher number of accidents, followed by age group 36-45 years.

### Research Question 2: Over time did the accidents decrease?


```{r}
group=factor(rep(1:5, each=40/5))
Accident_year<-ggplot(tidy_data, aes(x=Year_data)) + 
  geom_bar(fill=group,alpha=0.35, position="identity")+ ylab("Number of Accidents")+xlab("Years")
  labs(title = "Fig 2: Over time did the accidents decrease?", )
labels <- seq(1979,2018, length.out=10)
Accident_year+ scale_x_discrete(breaks=labels, "Years",)


```
Figure 2. shows the number of accidents over the years, the accidents were higher in 1983 and least in 2006. The accidents decrease of the period of time. 

### Research Question 3: Did higher speed limit zones have higher accidents?

```{r}

 ggplot(tidy_data, aes(Speed_limit,fill=Gender)) + 
  geom_bar() +xlim(15,70)+ylim(0,8)+theme_bw() +
  labs(title = "Fig 3: Did Higher speed limits lead to more accidents?", y = "Number of Accidents", x = "Year")+scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))

  
```
Figure 3. shows the distribution of accidents over speed limit with gender information, the accidents were when speed is higher and males have been in accidents more than others at lower speed limit.

### Research Question 4: Did Higher speed limits lead to severity in accidents?
```{r}
ggplot(joined_data, aes(Severity,Speed_limit,fill=Severity)) + geom_boxplot()+theme_bw()+labs(title = "Fig 4: Did Higher speed limits lead to severity in accidents?",x="Severity of the Accidents", y="Speed limit km/hr")+
ylim(10,70)+scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))

```

This plot has Severity on x-axis and speed limit on the road on y-axis. This plot shows that Fatal accidents happened at speed zones ranging from 20 km/hr to 60 km/hr with median of 30 km per hr, wheres as serious and slightly injured accidents happened at different speed limits. 



###  Research Question 4: Is there any correlation between speed, Number of Vehicles and Number of Casaulties ?
```{r}

mycor<-cor(select(tidy_data,Number_of_Vehicles,Number_of_Casualties,Speed_limit))
corrplot(mycor,method="number")



```
The correlation plot shows that there is some correlation between Number of Casualties and Number of Vehicles; no correlation between speed limit with Number of Vehicles and Number of Casualties.


### Conclusions

1. Research Questions 1: Younger age group (11-15 years) were involved in more accidents than the older group, this could be because of more younger population using bicycles compared to older age group. 

2. Accidents decreased over the years, which could be result of better laws for bicyclist and better road instructure.

3. Males had higher accident count even in lower speed limit zones.

4. Severity of the accident varied for different speed limits, fatal accidents happended in range 20-60 km/hr.


5. There was no significant correlation found between Number of Number of Casualties, Number of Vehicles and Speed limits.



## Bibliography 

[Bike Accidents in Great Britain](https://www.kaggle.com/datasets/johnharshith/bicycle-accidents-in-great-britain-1979-to-2018?select=Accidents.csv)


Rice, D., & Rolfe, M. (n.d.). Data Science Fundamentals: R Tutorials. Retrieved from https://classroom.google.com/u/1/w/MTIwNzU2NTkzNjI2/t/all














