---
title: "Statistics"
author: "Jaji Pamarthi"
date: "`r Sys.Date()`"
description: "Homework4"
output:
  distill::distill_article:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
    #code_folding: true

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE, cache=TRUE)
```

Bicyclist Accidents Data Analysis

## Packages used

```{r}
library(readr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Read Bikers information and Accident Data  

### Reading csv file and combining the two files

```{r}
Accident_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/Data/Accidents.csv")
Bikers_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/Data/Bikers.csv")


joined_data <- merge(Accident_data, Bikers_data, by.x = "Accident_Index", 
             by.y = "Accident_Index", all.x = TRUE, all.y = FALSE) # merging two csv files
head(joined_data)
```
### Tidy Data Generation
```{r}
joined_data<-joined_data%>%
 mutate(Year_data= format(Date, format="%Y"))

 #replace_na(joined_data,list(Road_conditions= "Missing Data",Weather_conditions="Missing Data"))


tidy_data<-select(joined_data, -Accident_Index, -Time, -Date, -Day,-Age_Grp)
tidy_data<-select(tidy_data,10,1:9)

colnames(tidy_data)

```


## Descriptive Statistics
### Mean, Median and Standard Deviation of Number of Vehicles involved in the accident and Number of Casualties
```{r}

summarize(joined_data,mean.Vehicles=mean(Number_of_Vehicles),median.Vehicles=median(Number_of_Vehicles),sd.Vehicles=sd(Number_of_Vehicles))

summarize(joined_data,mean.Casualties=mean(Number_of_Casualties),median.Casualties=median(Number_of_Casualties),sd.Casualties=sd(Number_of_Casualties))
```

Selecting only few columns to understand the changes in mean, median and standard deviation by group by Gender compared to mean,median and standard deviation calculated using all data.

```{r}

joined_data %>%
 select(Weather_conditions,Number_of_Vehicles,Number_of_Casualties) %>%
 group_by(Weather_conditions)%>%
 summarize(mean.Vehicles=mean(Number_of_Vehicles),median.Vehicles=median(Number_of_Vehicles),sd.Vehicles=sd(Number_of_Vehicles))

  
  

```
### Frequency based analysis for Categorical variables 

Table 1: Frequency of Gender

```{r}
joined_data %>%
  select(Gender) %>%
  table()
```
Table 2: Frequency of Severity of Accidents
```{r}
joined_data %>%
  select(Severity) %>%
  table()
```

## Plots


### Over time did the accidents decrease?


```{r}
 ggplot(joined_data, aes(Year_data)) + 
  geom_bar() + 
  labs(title = "Over time did the accidents decrease?")
```



### Did higher speed limit zones have higher accidents?
The plot below shows count of accidents for each speed limit. The accidents were highest at nominal speed limit that 30 km/hr. 
Limitations: The visualization is not clear about the units of the speed limit and y-axis needs to be scaled to make visualization readable. 

```{r}

 ggplot(joined_data, aes(Speed_limit)) + 
  geom_bar() + 
  labs(title = "Did Higher speed limits lead to more accidents?")+xlim(0,65)

  
```

### Bivariate Plot
```{r}
ggplot(joined_data, aes(Severity,Speed_limit)) + geom_boxplot()+labs(title = "Did Higher speed limits lead to severity in accidents?")+ylim(0,75)
```

This plot has Severity on x-axis and speed limit on the road on y-axis. This plot shows that Fatal accidents happened at speed zones ranging from 20 km/hr to 60 km/hr with mean of 30 km.hr, wheres as serious and slightly injured accidents happened at different speed limits. 
Limitations: Needs units about speed_limit 


##  Research Question yet to be anaswers


 Is there any correlation between Severity
and factors such as speed, road conditions, weather conditions, road type and light conditions?
