
---
title: "Better Visualizations"
author: "Jaji Pamarthi"
date: "`r Sys.Date()`"
description: "Homework5"
output: distill::distill_article
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE, cache=TRUE)
```

Bicyclist Accidents Data Analysis

## Required Libraries

```{r}
library(readr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(wesanderson)
library(corrplot)
```

## Read Bikers information and Accident Data  

### Reading csv file and combining the two files

```{r}
Accident_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/Data/Accidents.csv")
Bikers_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/Data/Bikers.csv")


joined_data <- merge(Accident_data, Bikers_data, by.x = "Accident_Index", 
             by.y = "Accident_Index", all.x = TRUE, all.y = FALSE) # merging two csv files
glimpse(joined_data)
```

### Tidy Data Generation

```{r}
joined_data<-joined_data%>%
 mutate(Year_data= format(Date, format="%Y"))


tidy_data<-select(joined_data, -Accident_Index, -Time, -Date, -Day,-Age_Grp)
tidy_data<-select(tidy_data,10,1:9)

glimpse(tidy_data)

```



## Visualizations


### Research Question 1: Over time did the accidents decrease?


```{r}
group=factor(rep(1:5, each=40/5))
Accident_year<-ggplot(joined_data, aes(x=Year_data)) + 
  geom_bar(fill=group) + xlab("Years")+ylab("Number of Accidents")+
  labs(title = "Fig 1: Over time did the accidents decrease?")
labels <- seq(1979,2018, length.out=10)
Accident_year+ scale_x_discrete(breaks=labels, labels<-as.character(labels))
plot(Accident_year)


ggplot(joined_data, aes(x=Year_data)) +geom_bar()+
      theme_bw() +
  labs(title = "Fig 2:Difference in Accidents per year for each Gender ", y = "Number of Accidents", x = "Year")+
  facet_wrap(vars(Gender), scales = "free")

ggplot(joined_data, aes(x=Year_data,fill = Gender)) +geom_bar()+
      theme_bw() +
  labs(title = "Fig 3: Difference in Accidents per year for each Gender ", y = "Number of Accidents", x = "Year")+
  facet_grid(vars(Weather_conditions), scales = "free")

```
Figure 1 shows the number of accidents over the years, the accidents were higher in 1983 and least in 2006. The accidents decrease of the period of time. Futher, Figure 2 shows the accidents for gender over 1979-2018 were the Males have higher accident counts. Suprisingly, most of the accidents took place in clear weather conditions according to Figure 3.


### Did higher speed limit zones have higher accidents?

```{r}

 ggplot(joined_data, aes(Speed_limit)) + 
  geom_bar()  + xlab("Speed Limit (km/hr)")+ylab("Number of Accidents")+
  labs(title = "Fig 4:Did Higher speed limits lead to more accidents?")+xlim(15,65)+ylim(0,8)

 ggplot(joined_data, aes(Speed_limit,fill=Gender)) + 
  geom_bar() +xlim(15,65)+ylim(0,8)+theme_bw() +
  labs(title = "Fig 5: Did Higher speed limits lead to more accidents?", y = "Number of Accidents", x = "Year")+scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))

  
```
Figure 4 shows the distribution of accidents over speed limit, the accidents were when speed is higher. Figure 5 shows the same Figure 4 with gender information, males have been in accidents more than others at lower speed limit.


## 
```{r}
ggplot(joined_data, aes(Severity,Speed_limit,fill=Severity)) + geom_boxplot()+theme_bw()+labs(title = "Fig 6: Did Higher speed limits lead to severity in accidents?",x="Severity of the Accidents", y="Speed limit km/hr")+
ylim(10,70)+scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))

```

This plot has Severity on x-axis and speed limit on the road on y-axis. This plot shows that Fatal accidents happened at speed zones ranging from 20 km/hr to 60 km/hr with mean of 30 km per hr, wheres as serious and slightly injured accidents happened at different speed limits. 



##  Research Question 3:Is there any correlation between speed, Number of Vehicles and Number of Casaulties ?
```{r}
mycor<-cor(select(tidy_data,Number_of_Vehicles,Number_of_Casualties,Speed_limit))
corrplot(mycor,method="number")


```
The correlation plot shows that there is some correlation between Number of Casualties and Number of Vehicles; no correlation between speed limit with Number of Vehicles and Number of Casualties.


## Reflections

The conclusions so far from the plots to the research are explained below each plot(Figure). 

I have all the information required for answering the research questions. 

For naive reader to understand the project better, the flow to the documents and information of the research questions is required. Along with the easily understandable plots. 

The plots generated have most of the information, I need to answer my research questions. But, some plots such as Figure 1, Figure 2 and Figure 3 need to be updated for better visualization.

Figure 2 and Figure 3 are not clear as the labels not readable and I would to work on updating the code for better visualization for next submission.







