---
title: "Homework1"
author: "Jaji Pamarthi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE, cache=TRUE)
```


# Reading Data from 

##Packages used

```{r}
library(readr)
library(magrittr)
library(dplyr)
```

## Read Austrailian Marriage Data 

Reading csv file
```{r}
Austrailian_marriage_data<-read_csv(file="C:/Users/jyoth/Documents/GitHub/Jajipamarthi/docs/Journal_files/Data/australian_marriage_tidy - australian_marriage_tidy.csv")

Austrailian_marriage_data
print(head(Austrailian_marriage_data))

```
## Data Type

There are three variable in this dataset: 

1. territory: It is contains names of the terriorty in Austraila. Data type is character 

2. resp : Represents response of the individual, binary (yes or no). Data type is charater

3. count: Population in the territory. Data type: Double precision integer.

4. percent : Percentage of poplation that expressed views. Data type: Double class float.

## Data Wrangling 

###Assigning new column names
```{r}
colnames(Austrailian_marriage_data) <- c("Territory_Name",
                       "Response_SameSexMarriage",
                       "Population",
                       "Percentage_Yes"
                       )

print(head(Austrailian_marriage_data))
```
### Using filter and arrange operations
```{r}

data(Austrailian_marriage_data)
Austrailian_marriage_data %>%
 filter(Percentage_Yes > 50) %>%
 arrange(desc(Percentage_Yes))
```
